---
title: "Data Collection and Data Wrangling"
author: "Uzair Shafique"
date: "Aug, 2023"
output: html_document
---

##### NOTE: 
1. Change author name and date to your exercise submission date in above section
2. Your code MUST execute without errors. 
3. You can add more lines in your code as required.

## Section 1: Manipulating Data

### Question 1 
**The dataset is loaded for you. Perform the following tasks:**  
1. use the USPersonalExpenditure dataset and store in a variable.  
2. Compute summary statistics: mean, standard deviation (sd) across time for each of the indicators.  
3. Create a new column that contains average values for each of the indicators.  


```{r }
library(help='datasets')
```


```{r }
USPersonalExpenditure
```

```{r }
### start solution ####
# Store the dataset in a variable
US_data <- USPersonalExpenditure
```


```{r }
#Compute summary statistics: mean and standard deviation across time for each indicator
summary(US_data)
```


```{r }
library(dplyr)
```


```{r }
# Create a new column with average values for each indicator
US_data$Average <- rowMeans(US_data)
US_data
```


```{r }
### end solution ####
```

### Question 2 
**download the data from the available URL:**  
1. Create a new column containing the average bodywt (body weight) of each genus and of each order.  
2. Create a dataframe containing the average REM sleep time for each order.  
3. How many missing values are there (in total) and per column?  
4. How would you like to impute missing values? Write Justification.  <i> Hint: Overall Mean/media/mode vs. Groupby Mean/media/mode?


```{r }
url <- "https://raw.githubusercontent.com/genomicsclass/dagdata/master/inst/extdata/msleep_ggplot2.csv"
filename <- "msleep_ggplot2.csv"
```


```{r }
### start solution ####
# Download and load the data
msleep_data <- read.csv(url)
```


```{r }
# Calculate average bodywt by genus and order
msleep_data <- msleep_data %>%
        group_by(genus) %>%
        mutate(avg_bodywt_genus = mean(bodywt, na.rm = TRUE)) %>%
        group_by(order) %>% 
        mutate(avg_bodywt_order = mean(bodywt, na.rm = TRUE))
```


```{r }
# Calculate average REM sleep time by order
avg_sleep_rem <- msleep_data %>%
  group_by(order) %>%
  summarize(avg_sleep_rem_time = mean(sleep_rem, na.rm = TRUE))
```


```{r }
# caluculate NA value's
sum(is.na(msleep_data))
# how to calculate missing values per column
missing_per_column <- colSums(is.na(msleep_data))
```


```{r }
# List of numeric columns
numeric_columns <- names(msleep_data)[sapply(msleep_data, is.numeric)]

# Impute missing values with column means for numeric columns
imputed_data <- msleep_data %>%
  mutate(across(all_of(numeric_columns), ~ ifelse(is.na(.), mean(., na.rm = TRUE), .)))
```


```{r }
### end solution ###
```

### Good job! You've completed this section!

## Section 2: Tidyverse

### Question 1
**Use the above dataset and perform the following tasks using any library from tidyverse:**  
1. Filter results to print average REM sleep and average total sleep  for those animals who are carnivores and then for those who are primates.  
2. Use the order column and "spread" it across the rest of the observations.  


```{r }
### start solution ####

```


```{r }
library(tidyverse)
library(dplyr)
library(ggplot2)
```


```{r }
# Task 1: Filter and print average REM sleep and average total sleep for carnivores and primates
msleep_data %>% filter(order == "Carnivora") %>% summarise(avg_rem = mean(sleep_rem), avg_sleep = mean(sleep_total))
```


```{r }
msleep_data %>% filter(order == "Primates") %>% summarise(avg_rem = mean(sleep_rem), avg_sleep = mean(sleep_total))
```


```{r }
# Task 2: Spread the order column across the rest of the observations
library(tidyr)
msleep_data %>% spread(order, name)
```


```{r }

```


```{r }

```


```{r }
### end solution ####
```

### Good job! You've completed this entire exercise!
